---
title: "MJMSTable"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MJMSTable}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `MJMSTable` package is designed to streamline the process of creating publication-ready statistical tables that adhere to the formatting requirements of the *Malaysian Journal of Medical Sciences* (MJMS). Built on top of `gtsummary` and `gt`, this package provides wrapper functions for common statistical tests, automatically applying the correct themes, layouts, and footnotes.

### Setup

First, load the package. For this vignette, we will use a hypothetical dataset called `mjms_data` that comes with the package.

```{r setup, eval=TRUE}
library(MJMSTable)

# View the structure of the dataset
head(mjms_data)

```

---

## 1. Descriptive Statistics

The `descriptive_tbl()` function provides a clean summary of demographic or baseline characteristics, grouped by a specific variable.

```{r descriptive, eval=TRUE}
descriptive_tbl(
  data = mjms_data,
  group_var = "Treatment_Group",
  included_var = c("Age", "BMI", "Sex", "Smoker"),
  non_normal_vars = c("Age"),
  table_caption = "Baseline Patient Characteristics",
  abbreviation = "BMI = Body Mass Index, SBP = Systolic Blood Pressure"
)

```

---

## 2. Comparing Means

### Independent t-test

Compare continuous variables between two independent groups using `ttest_tbl()`. You can toggle between standard and Welch's t-test using the `equal_var` argument.

```{r indep_ttest, eval=TRUE}
ttest_tbl(
  data = mjms_data,
  outcome_var = "BMI",
  group_var = "Sex",
  equal_var = TRUE,
  table_caption = "Comparison of BMI by Biological Sex",
  abbreviation = "BMI = body mass index"
)

```

### Paired t-test

For paired continuous data (e.g., pre- and post-measurements or paired body parts), use `paired_ttest_tbl()`.

```{r paired_ttest, eval=TRUE}
paired_ttest_tbl(
  data = mjms_data,
  id_var = "PatientID",
  pre_var = "SBP_Left_Arm",
  post_var = "SBP_Right_Arm",
  outcome_var = "SBP_Right_Arm",
  variable_label = "Systolic Blood Pressure (mmHg)",
  table_caption = "Comparison of SBP between Left and Right Arms"
)

```

### One-Way ANOVA

When comparing means across more than two groups, `anova_tbl()` provides the F-statistic, degrees of freedom, and post-hoc adjustment results. It can also generate an accompanying mean plot.

```{r anova, eval=TRUE}
# Returns a list containing both the table and a ggplot object
anova_results <- anova_tbl(
  data = mjms_data,
  outcome_var = "Cholesterol",
  group_var = "Treatment_Group",
  outcome_label = "Serum Cholesterol",
  show_plot = TRUE
)

# View the table
anova_results$table

# View the plot
anova_results$plot

```

---

## 3. Categorical Associations

### Chi-square Test

Evaluate the association between two categorical variables with `chisq_tbl()`.

```{r chisq, eval=TRUE}
chisq_tbl(
  data = mjms_data,
  exposure_var = "Sex",
  outcome_var = "Smoker",
  table_caption = "Association of Gender vs. Smoking Status"
)

```

### McNemar Test

For paired categorical data, `mcnemar_tbl()` computes McNemar's Chi-squared test with continuity correction.

```{r mcnemar, eval=TRUE}
mcnemar_tbl(
  data = mjms_data,
  pre_var = "Symptom_Pre",
  post_var = "Symptom_Post",
  table_caption = "Symptom status pre- and post-treatment",
  header_pre = "Pre Symptom",
  header_post = "Post Symptom"
)

```

---

## 4. Regression Analyses

### Linear Regression

Summarize Simple or Multiple Linear Regression models using `lm_tbl()`.

```{r lm, eval=TRUE}
lm_tbl(
  data = mjms_data,
  outcome_var = "Cholesterol",
  predictor_vars = c("Age", "BMI", "Smoker"),
  var_labels = list(Age ~ "Age (Years)", BMI ~ "Body Mass Index"),
  table_caption = "Multiple Linear Regression for Cholesterol"
)

```

It also supports interaction, which can be specified using "*" or ":" (similar to lm or glm formula):
```{r lm_x, eval=TRUE}
lm_tbl(
  data = mjms_data,
  outcome_var = "Cholesterol",
  predictor_vars = c("Age", "BMI*Smoker"), # or predictor_vars = c("Age", "BMI", "Smoker", "BMI:Smoker"),
  var_labels = list(Age ~ "Age (Years)", BMI ~ "Body Mass Index"),
  table_caption = "Multiple Linear Regression for Cholesterol"
)
```

### Logistic Regression

For binary outcomes, `logistic_tbl()` outputs Adjusted Odds Ratios, 95% CIs, and Wald statistics. You can also explicitly define reference levels.

```{r logistic, eval=TRUE}
logistic_tbl(
  data = mjms_data,
  outcome_var = "Heart_Disease",
  predictor_vars = c("Age", "Sex", "Smoker"),
  ref_levels = list(Smoker = "No", Sex = "Female"),
  cat_vars = c("Sex", "Smoker"),
  table_caption = "Risk Factors for Heart Disease"
)

```

It also supports interaction, which can be specified using "*" or ":" (similar to lm or glm formula):
```{r logistic_x, eval=TRUE}
logistic_tbl(
  data = mjms_data,
  outcome_var = "Heart_Disease",
  predictor_vars = c("Age", "Sex", "Smoker", "Age:Smoker"), # or predictor_vars = c("Sex", "Age*Smoker"),
  ref_levels = list(Smoker = "No", Sex = "Female"),
  cat_vars = c("Sex", "Smoker"),
  table_caption = "Risk Factors for Heart Disease"
)
```

---

## 5. Advanced Analyses

### Diagnostic Test (Sensitivity, Specificity, AUC)

Analyze the performance of medical markers against a gold standard using `diagnostic_tbl()`. It handles multiple markers at once and optionally generates an ROC plot.

```{r diagnostic, eval=TRUE}
# Define index markers and their cutoffs
med_marker <- list(
  Cholesterol  = list(label = "Serum Cholesterol (Cutoff: 200)", cutoff = 200),
  BMI          = list(label = "Body Mass Index (Cutoff: 25)", cutoff = 25),
  SBP_Left_Arm = list(label = "Systolic BP - Left (Cutoff: 140)", cutoff = 140)
)

diag_results <- diagnostic_tbl(
  data = mjms_data,
  status_var = "Gold_Standard",
  marker_map = med_marker,
  show_plot = TRUE,
  plot_marker = "Cholesterol",
  table_caption = "Diagnostic Accuracy of Cardiovascular Markers",
  abbreviation = "AUC = Area Under Curve; SBP = Systolic Blood Pressure"
)

# View outputs
diag_results$table
diag_results$plot

```

### Pearson Correlation

Generate a clean correlation matrix with embedded p-values and standard deviations using `pearson_tbl()`. A scatterplot with an LM trendline can be toggled on.

```{r pearson, eval=TRUE}
corr_results <- pearson_tbl(
  data = mjms_data,
  included_var = c("Age", "BMI", "SBP_Left_Arm", "SBP_Right_Arm"),
  show_plot = TRUE,
  plot_x = "SBP_Left_Arm",
  plot_y = "SBP_Right_Arm",
  abbreviation = "BMI = body mass index; SBP = systolic blood pressure",
  table_caption = "Correlation between clinical variables"
)

# View outputs
corr_results$table
corr_results$plot

```


